### 2019-07-20 1_introduction.tex
revised intro

There is a second important issue: How does one predict how
the performance of a mini-app generalizes to a full
application, even one closely related? Part of the problem is
that \emph{weak} and \emph{strong} scaling, by far the most commonly
reported performance metrics,
do not directly characterize the performance of the underlying
programming system. Weak scaling can hide arbitrary amounts of runtime
system overhead with sufficiently large problem sizes, and strong
scaling does not separate runtime system overhead from costs
(such as increased communication) that scale with the number of nodes when
using progressively larger portions of a machine.

To address the first issue, we present Task Bench, a
\emph{parameterized} benchmark designed for exploring the performance
behaviors of parallel and distributed programming systems under a
variety of conditions. A critical property of Task Bench is that,
while able to express a wide variety of typical parallel programming
patterns, it is simple enough to be implemented with reasonable effort
to a level of quality sufficient for comparative
studies. Task Bench provides a core API that is shared across the
various system implementations. New benchmarks created with Task Bench
immediately run on all systems, and new systems immediately run all
benchmarks. Thus, Task Bench is much more productive for benchmarking
a wide range of phenomena than the standard approach in which each
benchmark must be implemented separately for each system.

A key insight of Task Bench is that many applications can be modeled
as a set of \emph{tasks}, or coarse-grain units of work, with
dependencies between tasks representing the communication and synchronization required for
distributed execution. Task Bench consists of a configurable \emph{task graph} with tasks for
each point in an iteration space. Tasks are connected by dependencies
as determined by a \emph{dependence relation}. This design permits a
very concise description of a wide variety of patterns relevant to
scientific computing and large scale data analysis: trivial parallelism, halo exchanges (such as
seen in structured and unstructured mesh codes), sweeps (such as used
in the discrete ordinates method of radiation simulation), FFTs, trees
(for divide and conquer algorithms), and so on. The tasks at each
point in the task graph can be configured to execute kernels with a
variety of computational properties, including compute- and
memory-bound operations of varying durations. The dependencies between
tasks can be configured to carry communication payloads of varying size, permitting
the design of communication-bound cases. Finally, multiple
(potentially heterogeneous) task graphs can be executed concurrently
to introduce task parallelism into the workload. Together, these
design elements permit the exploration of a large space of application
behaviors.

To address the second issue and characterize the contribution of a programming model's runtime system to application performance, we introduce a new metric called \emph{minimum effective task granularity} (METG). Intuitively, METG(50\%) is the smallest executable task granularity that maintains at least 50\% efficiency, meaning that the application achieves at least 50\% of peak
performance (e.g., FLOP/s or memory bandwidth) on a given machine. METG captures the essence of what is important in a weak
or strong scaling study, the behavior at the limit of scalability. For
weak scaling, METG(50\%) corresponds directly to the smallest problem size that can be weak-scaled with 50\% efficiency. For strong scaling, METG(50\%) can be used to compute the scale at which performance can be expected to dip below 50\% efficiency.

A lower METG does not necessarily mean that performance for a particular workload is
better. For example, two systems with METG(50\%) of 100~\textmu{}s and 1~ms,
respectively, running an application with an average task granularity
of 10~ms, are both very likely to perform well. Only when the task
granularity approaches (or drops below) METG(50\%) will performance
likely diverge. Thus METG helps to identify the regime in which a
given system can deliver good performance.

Task Bench and METG address issues common in limit studies of runtime
systems for parallel and distributed programming. Such studies often
employ the metric \emph{tasks per second} (TPS), which is almost
universally measured with trivial (i.e., no) dependencies \cite{LegionTracing18, Canary16, Armstrong14}. While
phrased in terms of tasks, TPS can be measured for any system as long
as the application in question has identifiable units of work that run
to completion without interruption. TPS is an upper bound on
runtime-limited application throughput. But it is not a tight bound, as the
cost of non-trivial dependencies can be significant. This issue can be easily fixed by running non-trivial
configurations of Task Bench.

@@

\begin{itemize}
\item Section~\ref{sec:metg} defines the METG(50\%) metric and its
  relationship to quantities of interest to application developers.
\item Section~\ref{sec:task-bench} describes the design of Task Bench.
\item Section~\ref{sec:implementation} details the implementation of
  Task Bench in 15 programming systems.
\item Section~\ref{sec:evaluation} provides a comprehensive evaluation
  of Task Bench on up to 256 Haswell nodes of the Cori
  supercomputer~\cite{Cori}.
\end{itemize}

>>>>

We present Task Bench, a benchmark designed for exploring the performance
behaviors of parallel and distributed programming systems under a
variety of conditions.  The key property of TaskBench is that it completely separates
the implementation of TaskBench for a new system to be measured from the implementation
of a new benchmark application to be run on the systems supported by TaskBench.
In all benchmarks we know of, the effort to implement $n$ benchmarks on $m$
systems is ${\cal O}(mn)$.  TaskBench's design reduce the work to ${\cal O}(m) + {\cal O}(n)$,
enbaling dramatically more systems and benchmarks to be explored for the same amount of programming
effort.  New benchmarks created with Task Bench
immediately run on all systems, and new systems immediately run all
benchmarks.

Benchmarks in Task Bench are based on the observation that many applications can be modeled
as a set of \emph{tasks}, or coarse-grain units of work, with
dependencies between tasks representing the communication and synchronization required for
distributed execution. A benchmark consists of a parameterized \emph{task graph} with tasks for
each point in an iteration space. Tasks are connected by dependencies
as determined by a \emph{dependence relation}. This design permits a
very concise description of a wide variety of patterns relevant to
scientific computing and large scale data analysis: trivial parallelism, halo exchanges (such as
seen in structured and unstructured mesh codes), sweeps (such as used
in the discrete ordinates method of radiation simulation), FFTs, trees
(for divide and conquer algorithms), and so on. The tasks at each
point in the task graph can be configured to execute kernels with a
variety of computational properties, including compute- and
memory-bound operations of varying durations. The dependencies between
tasks can be configured to carry communication payloads of varying size, permitting
the design of communication-bound cases. Finally, multiple
(potentially heterogeneous) task graphs can be executed concurrently
to introduce task parallelism into the workload. Together, these
design elements permit the exploration of a large space of application
behaviors.

Adding a new programming system to Task Bench involves  implementing a set of standard
services, such as executing a task or a data transfer.  This interface is simple enough to 
be implemented with reasonable effort to a level of quality sufficient for comparative
studies. Task Bench also provides a core library that encapsulates the functionality
shared by all system implementations, which not only reduces the implementation effot
required but also makes it much easier to achieve truly apples-to-apples comparisons
between systems.

Using Task Bench we were able to quickly incorporate 13 very different parallel and distributed
runtime systems.  By running all systems on common benchmarks we were able to quantify phenomena
that have never before been measured.
Most strikingly, the performance of the various systems we examine varies by more than five orders
of magnitude, with very popular, widely used systems at both ends of the spectrum!  Clearly
the slower systems have ``good enough'' performance for some class of applications, while presumably
providing advantages in programmer productivity.

These results point to a natural question:   How does one predict whether the performance of a runtime
system will in fact be good enough for a particular application?  In trying to answer this question,
we realized that \emph{weak} and \emph{strong} scaling, by far the most commonly
reported performance metrics, do not directly characterize the performance of the underlying
programming system. Weak scaling can hide arbitrary amounts of runtime
system overhead by using sufficiently large problem sizes, and strong
scaling does not separate runtime system overhead from costs
(such as increased communication) that scale with the number of nodes when
using progressively larger portions of a machine.

To characterize the contribution of a programming model's runtime
system to application performance, and as an example of the novel studies that can be done
with Task Bench, we introduce a new metric called
\emph{minimum effective task granularity} (METG). Intuitively, for a given
workload, METG(50\%) is the smallest task granularity that maintains
at least 50\% efficiency, meaning that the application achieves at
least 50\% of peak performance (e.g., FLOP/s or memory bandwidth) on a
given machine. METG captures the essence of what is important in a
weak or strong scaling study, the behavior at the limit of
scalability. For weak scaling, METG(50\%) corresponds directly to the
smallest problem size that can be weak-scaled with 50\%
efficiency. For strong scaling, METG(50\%) can be used to compute the
scale at which performance can be expected to dip below 50\%
efficiency.  We note that METG(50\%) for a given runtime system will
vary with the particular application and the underlying hardware---i.e., METG(50%)
is not a constant for a given system, but we will see that individual systems have
a characteristic ranges of METG(50\%) values and that there is additional insight
in the reasons that the mininum effective task granularity can vary.

Another important characteristic is that a lower METG does not necessarily mean that 
performance for a particular workload is better. For example, two systems with METG(50\%) of 100~\textmu{}s and 1~ms,
respectively, running an application with an average task granularity
of 10~ms, are both very likely to perform well. Only when the task
granularity approaches (or drops below) METG(50\%) will performance
likely diverge. Thus METG helps to identify the regime in which a
given system can deliver good performance, and helps to explain how
different systems coexist with METG(50\%) values that vary by orders of magnitude.

%
% FIXME:  Move these two paragraphs out of the introduction and into one of the METG sections.
%
Task Bench and METG address issues common in limit studies of runtime
systems for parallel and distributed programming. Such studies often
employ the metric \emph{tasks per second} (TPS), which is almost
universally measured with trivial (i.e., no) dependencies \cite{LegionTracing18, Canary16, Armstrong14}. While
phrased in terms of tasks, TPS can be measured for any system as long
as the application in question has identifiable units of work that run
to completion without interruption. TPS is an upper bound on
runtime-limited application throughput. But it is not a tight bound, as the
cost of non-trivial dependencies can be significant. This issue can be easily fixed by running non-trivial
configurations of Task Bench.

@@

\begin{itemize}
\item Section~\ref{sec:task-bench} describes the design of Task Bench.
\item Section~\ref{sec:implementation} details the implementation of
  Task Bench in 15 programming systems.
\item Section~\ref{sec:metg} defines the METG(50\%) metric and its
  relationship to quantities of interest to application developers.
\item Section~\ref{sec:evaluation} provides a comprehensive evaluation
  of Task Bench on up to 256 Haswell nodes of the Cori
  supercomputer~\cite{Cori}.
\end{itemize}
